### 2015-05-29 reports/paper/conclusion.tex
converted to thesis format

\section{Conclusion}
In this paper, we proposed and built an inexpensive, portable yet reasonably accurate sound localization system with two microphone arrays. We analyzed different array architectures and demonstrated that a two array architecture achieves better accuracy compared to a single array system of similar size. By using cross-correlation output as the likelihood for different arrival time differences, each array generates a likelihood map for each possible location in the area. We demonstrated that by merging likelihood maps from the two arrays, we were able to achieve good localization accuracy (less than $3$ cm average error) in a local one meter by one meter region.

This system can be used in HCI applications that incorporate sound position and movement information, such as drawing with music, or can be used to create AI games with physical pieces where the computer controls some of the pieces. This system can also be used to in Augmented Reality (AR) applications that allow user computer interaction with music tags.

For future directions, this system can be extended with another microphone on each array to perform 3D localization. Another possible direction is to investigate multi-source localization using digitally tagged audio sources.